---
title: "EDA"
Name: "Tyusha Sarawagi"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mosaic)
library(Stat2Data)
library(ggplot2)
library(tidyverse)
library(moments)
library(imputeTS)

data <- read.csv("~/countries of the world.csv")

```


### Data Cleaning

Getting rid of the column with missing data is probably not a good idea because by doing so, I lost the data for 48 countries. 

Currently we have 110 missing values.


```{r}
sum(is.na(data))
```

The columns Industry and Agriculture had a lot of missing values and I was not going to use these variable for my model, so I decided to get ride of them. 


```{r}

cols.dont.want <- c("Agriculture", "Industry")
data2 <- data[, ! names(data) %in% cols.dont.want, drop = F]

sum(is.na(data2))

```

Before, we had 111 missing values but after removing we have 79 which is a good progress. 


I looked at the dataset and rows West Sahara and Walis and Futuna has a lot of missing values so I am going to get rid of them.
```{r}
data3 = data2[-c(222), ]

data4 = data3[-c(223), ]

sum(is.na(data4))
```

Now, we have 67 missing values in our dataset which are randomly scattered throughout the dataset so I am going to substitute missing cells with column mean. 

```{r}

#library(zoo)
#na.aggregate(data4)

#data6<- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
#data4[] <- lapply(data4, data6)

data6 = na_mean(data4)

```


We can see that all the missing values have been taken care of at this point. 

```{r}
sum(is.na(data6))

```


The main variables that I want to use are Net Migration, GDP per capita, Literacy, Population density, Death Rate, Service, Infant Mortality, Phones, and Region. 



### Plots and Summary Statistics



```{r}
table(data6$Region)
```

```{r}
favstats(data6$Net.migration)
```

```{r}
favstats(data6$Pop..Density..per.sq..mi..)
```

```{r}
favstats(data6$GDP....per.capita.)
```

```{r}
favstats(data6$Literacy....)
```

```{r}
favstats(data6$Deathrate)
```

```{r}
favstats(data6$Service)
```


```{r}
favstats(data6$Infant.mortality..per.1000.births.)
```

```{r}
favstats(data6$Phones..per.1000.)
```



```{r}
gf_point(Net.migration~ GDP....per.capita., data=data6)
```

```{r}
gf_point(Net.migration~ Literacy...., data=data6)
```


```{r}
gf_point(Net.migration ~ Pop..Density..per.sq..mi.., data=data6)

```



```{r}
gf_point(Net.migration ~ Infant.mortality..per.1000.births., data=data6)
```



```{r}
gf_point(Net.migration ~ Phones..per.1000. , data=data6)
```


```{r}
gf_point(data6$Net.migration~ data6$Deathrate)
```



```{r}
gf_point(data6$Net.migration~ data6$Service)
```




```{r fig.width=5, fig.height=5 }

#gf_boxplot(Net.migration~ Region, data = data6, las=2, par(mar = c(12, 5, 4, 2)+ 0.1))


library(ggplot2)
ggplot(data6, aes(Region, Net.migration)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))


```


I decided to see the migration rate for each region and which region has the higest and lowest migration rate. 

```{r}
data7 = data6 %>% group_by(Region) %>% summarise(new_migration =mean(Net.migration))
data7


ggplot(data7, aes(Region, new_migration)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))

```


### Correlation between the Variables 

I am going to find the correlation between Net Migration and all the other variables to find out what variables should I add in my model. 

```{r width=2, fig.height=2}

cor(~cbind(Net.migration, Literacy...., Deathrate, Service),data=data6)

```


```{r}
cor(~cbind(Net.migration,GDP....per.capita., Phones..per.1000.),data=data6)
```


```{r}
data7 = rename(data6, Infant.mortality= Infant.mortality..per.1000.births., Pop..Density=Pop..Density..per.sq..mi..)

cor(~cbind(Net.migration,Infant.mortality, Pop..Density),data=data7)
```



